\tool{}: Specification- and Syntax-based Automated Testing Linguistic Capabilities of NLP Models

### 2023-01-18 paper/fse23/experiment.tex
debug

\begin{enumerate}[label=\textbf{RQ\arabic*}]
\item \label{rq:three}: \textbf{Test Case Diversity.} Can \tool generate more diverse test cases
  than \Cklst?
\item \label{rq:one}: \textbf{Test Oracle.}  Can \tool generate test sentences with correct sentiment labels?
\item \label{rq:two}: \textbf{Capability Testing.} Are \tool generated test sentences correctly categorized into a \lc?
% \item \label{rq:four}: \textbf{Usefulness of Test Inputs.} Can \tool be useful to find root causes of bugs
%   in the \sa models?
%% \item \label{rq:three}: How effective is our new test case generation
%%   using \cfg expansion? % ablation study
\end{enumerate}

@@

\subsection{Experimental Subjects}

\paragraph*{\textbf{NLP Models \& Dataset}}
We evaluate our approach on three learning-based \sa models released from the Hugging Face centralized
model hub:\footnote{https://huggingface.co/models} \texttt{\Bert}
(\bertsamodel), \texttt{\Roberta} (\robertasamodel), and \texttt{\Dbert} (\disbertsamodel). These models were pre-trained on English
language using a masked language modeling (MLM) objective, and
were fine-tuned on the \sa task. In this experiment, we used the \Sst \cite{socher2013sst} dataset for
searching the seeds in \tool. \Sst is a
corpus of movie review, consisting of 11,855 sentences
with sentiment scores. We split the
scores into ranges [0, 0.4], (0.4, 0.6] and (0.6, 1.0] to assign them
negative, neutral and positive labels, respectively. In addition,
\Swn~\cite{baccianella2010sentiwordnet} is a publicly available dataset for English sentiment lexicons.  
It provides lexical sentiment scores and the sentiment word labels 
are categorized by implementing the rules.
We used \Swn as both the generation and the expansion domain knowledge as shown in Figure \ref{fig:overview}.

\paragraph*{\textbf{Comparison Baseline}}
We compared \tool with
\Cklst,\footnote{https://github.com/marcotcr/checklist} a
manual template and \lc based approach to generate test cases. In this evalaution, we used \Cklst's \sa test cases that are
generated from its publicly available Jupyter Notebook implementation.

@@

\paragraph*{\textbf{RQ1}}
Recall that \Cklst relies on significant manual efforts and may not generate comprehensive test cases in a linguistic capability. \tool, instead, automatically generates test cases based on a search
dataset and the syntax in a large reference corpus. 
We expect \tool can generate a more diverse test suite than \Cklst.

We first evaluate the three \sa models by testing them on the \tool  test cases, reporting number of test cases and each model's failure rate in each linguistic capability. Specifically, for each \lc, we randomly choose 50, 100 and 200 seeds , and expand test cases from these seeds. We selected subsets of seeds to be used for expansion because the syntax-based expansion phase may be expensive to run all the generated seeds. We tested the three models using both seed and expanded test cases. To account for the randomness in random seed selection, we repeated these experiments 3 times and report median number of test cases and each model's failure rate.

@@

\paragraph*{\selfbleu} We reuse the input diversity metric, 
called \selfbleu, that Zhu \etal introduced~\cite{zhu2018texygen}. \bleu evaluates the token-level similarity.
% Regarding that \selfbleu takes each sentence as hypothesis and rest in a collection of textual data as reference, we calculate \bleu scores for every pairs of hypothesis and each reference sentence.
The \selfbleu is defined as the average \bleu scores over all reference sentences.
% Since the \bleu score ranges from 0 as the least similar inputs to 1 as the most similar inputs,
A higher \selfbleu score indicates less diversity in the test suite. In the experiment, we collected 50, 100 and 200 randomly selected \tool seeds and reported the median \selfbleu score over three trials for each group of seeds.
% The median \selfbleu over 3 trials of the random collection of the 50, 100 and 200 seeds is defined as the final \selfbleu scores. For the comparison, \selfbleu scores of \Cklst test cases are calculated in the same manner.

\paragraph*{Production rule coverage.} We propose a new metric to evaluate the syntactic diversity of the generated test suite.
It is defined as the number of production rules used in a set of test sentences. In our experiments, we used the Berkeley Neural Parser~\cite{kitaev2018seedparser,kitaev2019seedparser} to parse and collect all the production covered in a set of test sentences.
We compared the \pdr between 50, 100 and 200 randomly selected \tool seeds and the \Cklst test cases.
% collect the  of 50, 100 and 200 randomly selected \tool seeds and \Cklst test cases by the Berkeley Neural Parse~\cite{kitaev2018seedparser,kitaev2019seedparser}. Next, we traverse the parse trees and the \pdr score is calculated as the number of the production rules for the 50, 100 and 200 selected \tool seeds and \Cklst test cases.

@@

% ($Cov(\mathcal{X})$ in \equref{eq:coverage}),
% where $N$ is the total number blocks, $\mathbb I(\cdot)$ is the indicator function, and
% $(B_i(x) > \tau_i))$ represents whether $i^{th}$ block is activated by input $x$~(the definition of $B_i$ and $\tau_i$ are the same with \equref{eq:new2} and \equref{eq:new3}).
% Because AdNNs activate different blocks for decision making, then a higher block coverage indicates the test samples cover more decision behaviors.

@@

% For each subject, we randomly select 100 seed samples from the test dataset as seed inputs. We then feed the same seed inputs into \tool and \texttt{ILFO} to generate test samples. 
% Finally, we feed the generated test samples to AdNNs and measure block coverage.
% We repeat this process 10 times and record the average coverage and the variance.
% The results are shown in \tabref{tab:coverage} last two columns.

@@

\paragraph*{\textbf{RQ2 and RQ3}} As described in Section \ref{sec:approach},
\tool generates test cases in two steps: specification-based seed
generation and syntax-based sentence expansion. These automated steps
may generate seed/expanded sentences marked with incorrect sentiment
labels or categorized into wrong linguistic capabilities.
% For example,
% the search rule and template defined in a linguistic capability may
% not always generate seed sentences in that capability or with the
% correct label. 
To answer RQ2 and RQ3, we performed a manual study to
measure the correctness of the sentiment labels and linguistic
capabilities associated with the seed/expanded sentences, produced by
\tool.

>>>>

\begin{enumerate}[label=\textbf{RQ\arabic*}]
% \item \label{rq:three}: \textbf{Test Case Diversity.} Can \tool generate more diverse test cases
%   than \Cklst?
\item \label{rq:three}: \textbf{Test Efficiency.} Can \tool generated test cases evaluate an NLP model more efficiently?
\item \label{rq:one}: \textbf{Test Oracle.}  Can \tool generate test sentences with correct sentiment labels?
\item \label{rq:two}: \textbf{Capability Testing.} Are \tool generated test sentences correctly categorized into a \lc?
% \item \label{rq:four}: \textbf{Usefulness of Test Inputs.} Can \tool be useful to find root causes of bugs
%   in the \sa models?
%% \item \label{rq:three}: How effective is our new test case generation
%%   using \cfg expansion? % ablation study
\end{enumerate}

@@

\subsection{Experimental Subjects}
\paragraph*{\textbf{NLP Models \& Dataset}}
We evaluate our approach on three \sa models and two \hsd models. All
are learning-based models released from the Hugging Face centralized
model hub:\footnote{https://huggingface.co/models} \texttt{\Bert}
(\bertsamodel), \texttt{\Roberta} (\robertasamodel), and
\texttt{\Dbert} (\disbertsamodel~\cite{sanh2019distilbertad}) for \sa
task, and \texttt{\hsdBert} (\berthsdmodel~\cite{aluru2020deep}) and
\texttt{\hsdRoBerta} (\robertahsdmodel) for \hsd.

These models were pre-trained on English language using a masked
language modeling (MLM) objective, and were fine-tuned on the \sa and
\hsd task. In this experiment, we used the \Sst~\cite{socher2013sst}
and \Hatexp~\cite{mathew2021hatexplain} datasets, for \sa and \hsd
respectively, for searching the seeds in \tool. \Sst is a corpus of
movie review, consisting of 11,855 sentences with sentiment scores. We
split the scores into ranges [0, 0.4], (0.4, 0.6] and (0.6, 1.0] to
    assign them negative, neutral and positive labels,
    respectively. \Hatexp is a hate speech dataset collected from
    Twitter and Gab. It consists of 20,148 sentences (9,055 from
    Twitter and 11,093 from Gab). Each data has sentence, class of
    hate speech, target community, and rationale for the labelling
    decision.  In addition, \Swn~\cite{baccianella2010sentiwordnet} is
    a publicly available dataset for English sentiment lexicons.  We
    used \Swn as both the generation and the expansion domain
    knowledge as shown in Figure \ref{fig:overview}.

\paragraph*{\textbf{Comparison Baseline}}
We compared \tool with
\Cklst,\footnote{https://github.com/marcotcr/checklist} a manual
template and \lc based approach to generate test cases. In this
evalaution, we used \Cklst's \sa test cases that are generated from
its publicly available Jupyter Notebook implementation. \jl{Do we need
  to explain why we used checklist only as a baseline?}

@@

\paragraph*{\textbf{RQ1}}
Recall that \Cklst relies on significant manual efforts and may not
generate comprehensive test cases in a linguistic capability. \tool,
instead, automatically generates test cases based on a search dataset
and the syntax in a large reference corpus.  We expect \tool can
generate more efficient test suite than \Cklst for finding errors with
regards to quantity and diversity of test suite.

We first evaluate the three \sa models and two \hsd models by testing
them on the \tool test cases, reporting number of test cases and each
model's failure rate in each linguistic capability. Specifically, for
each \lc, we expand test cases from all generated seeds.
% We selected subsets of seeds to be used for expansion because the
% syntax-based expansion phase may be expensive to run all the
% generated seeds.
We test the five models using both seed and expanded test cases. 
% To account for the randomness in random seed selection, we repeated
% these experiments 3 times and report median number of test cases and
% each model's failure rate.

@@

\paragraph*{\selfbleu} We reuse the input diversity metric, 
called \selfbleu, that Zhu \etal
introduced~\cite{zhu2018texygen}. \bleu evaluates the token-level
similarity.
% Regarding that \selfbleu takes each sentence as hypothesis and rest
% in a collection of textual data as reference, we calculate \bleu
% scores for every pairs of hypothesis and each reference sentence.
The \selfbleu is defined as the average \bleu scores over all
reference sentences.
% Since the \bleu score ranges from 0 as the least similar inputs to 1
% as the most similar inputs,
A higher \selfbleu score indicates less diversity in the test
suite. In the experiment, we collected 50, 100 and 200 randomly
selected \tool seeds and reported the median \selfbleu score over
three trials for each group of seeds.
% The median \selfbleu over 3 trials of the random collection of the
% 50, 100 and 200 seeds is defined as the final \selfbleu scores. For
% the comparison, \selfbleu scores of \Cklst test cases are calculated
% in the same manner.

\paragraph*{Production rule coverage.} We propose a new metric to evaluate the syntactic diversity of the generated test suite.
It is defined as the number of production rules used in a set of test
sentences. In our experiments, we used the Berkeley Neural
Parser~\cite{kitaev2018seedparser,kitaev2019seedparser} to parse and
collect all the production covered in a set of test sentences.  We
compared the \pdr between 50, 100 and 200 randomly selected \tool
seeds and the \Cklst test cases.
% collect the of 50, 100 and 200 randomly selected \tool seeds and
% \Cklst test cases by the Berkeley Neural
% Parse~\cite{kitaev2018seedparser,kitaev2019seedparser}. Next, we
% traverse the parse trees and the \pdr score is calculated as the
% number of the production rules for the 50, 100 and 200 selected
% \tool seeds and \Cklst test cases.

@@

% ($Cov(\mathcal{X})$ in \equref{eq:coverage}), where $N$ is the total
% number blocks, $\mathbb I(\cdot)$ is the indicator function, and
% $(B_i(x) > \tau_i))$ represents whether $i^{th}$ block is activated
% by input $x$~(the definition of $B_i$ and $\tau_i$ are the same with
% \equref{eq:new2} and \equref{eq:new3}).  Because AdNNs activate
% different blocks for decision making, then a higher block coverage
% indicates the test samples cover more decision behaviors.

@@

% For each subject, we randomly select 100 seed samples from the test
% dataset as seed inputs. We then feed the same seed inputs into \tool
% and \texttt{ILFO} to generate test samples.  Finally, we feed the
% generated test samples to AdNNs and measure block coverage.  We
% repeat this process 10 times and record the average coverage and the
% variance.  The results are shown in \tabref{tab:coverage} last two
% columns.

@@

\paragraph*{\textbf{RQ2 and RQ3}} \jl{need to do the manual study for the \hsd task also?} As described in Section \ref{sec:approach},
\tool generates test cases in two steps: specification-based seed
generation and syntax-based sentence expansion. These automated steps
may generate seed/expanded sentences marked with incorrect sentiment
labels or categorized into wrong linguistic capabilities.
% For example,
% the search rule and template defined in a linguistic capability may
% not always generate seed sentences in that capability or with the
% correct label. 
To answer RQ2 and RQ3, we performed a manual study to measure the
correctness of the sentiment labels and linguistic capabilities
associated with the seed/expanded sentences, produced by \tool.
